---
title: "ADS 500B Project"
author: "Madeline Chang"
date: "2024-03-01"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
library(tidymodels)
library(mosaic)
library(cluster)   
library(factoextra)
library(lubridate)
```

```{r}
housing_sales<- read.csv("~/ADS/ADS 500B/Datasets/house_sales.csv") #reading in data
```

```{r}
housing_sales[2] <- housing_sales[2] %>%
  mutate(date = as.character(date)) %>%
  mutate(date = str_remove(date, "T000000")) %>%
  mutate(date = ymd(date))

str(housing_sales)
```

```{r}
house_na<- housing_sales %>%
  filter(is.na(housing_sales[,1]) | is.na(housing_sales[,2]) | is.na(housing_sales[,3]) | is.na(housing_sales[,4]) | is.na(housing_sales[,5]) | is.na(housing_sales[,6]) | is.na(housing_sales[,7]) | is.na(housing_sales[,8]) | is.na(housing_sales[,9]) | is.na(housing_sales[,10])| is.na(housing_sales[,11]) | is.na(housing_sales[,12]) | is.na(housing_sales[,13]) | is.na(housing_sales[,14]) | is.na(housing_sales[,15]) | is.na(housing_sales[,16]) | is.na(housing_sales[,17]) | is.na(housing_sales[,18]) | is.na(housing_sales[,19]) | is.na(housing_sales[,20]) | is.na(housing_sales[,21]))
```

```{r}
housing_clean<- housing_sales %>%
  filter(price != is.na(price))
```

Choosing how to clean the NAs out kind of relies on what variables we want to focus on. From the data I've seen, I think we could do a supervised learning with an equation that looks like the following:

price = intercept + var1 + var2 + var3 + var4 + etc.

I think I definitely want to include yr_built, zipcode, sqft_living, sqft_lot, bedrooms, and bathrooms in our exploratory data analysis. I would also suggest that we look at the condition, just to see how things line up.

# Notes for Bedrooms (1134 NAs)

```{r}
ggplot(data = housing_clean) +
  geom_boxplot(aes(y=bedrooms))
```

- the row with 33 bedrooms should be smoothed (only sold for 640K)
- replace with median due to skew? or just remove rows with NAs


# Notes for Bathrooms (1068 NAs)

```{r}
ggplot(data = housing_clean) +
  geom_boxplot(aes(y=bathrooms))
```

- replace with median due to skew? or just remove rows with NAs


# Notes for Square Feet of Living Space (1110 NAs)

```{r}
ggplot(data = housing_clean) +
  geom_boxplot(aes(y=sqft_living))
```

- Remove rows with NAs?

# Notes for Square Feet of Lot Space (1044)

```{r}
ggplot(data = housing_clean) +
  geom_boxplot(aes(y=sqft_lot))
```

- replace with median due to skew? or just remove rows with NAs
- sqft lot > 1500000 -> nosiy?

# Notes for Year Built

```{r}
ggplot(data = housing_clean) +
  geom_bar(aes(x=yr_built))
```

- No NAs

# Notes for Year Condition

```{r}
ggplot(data = housing_clean) +
  geom_bar(aes(x=condition))
```

- No NAs
