---
title: "ADS 500B Project"
author: "Madeline Chang"
date: "2024-03-01"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
library(dplyr)
library(tidyverse) 
library(tidymodels)
library(mosaic) 
library(cluster)  
library(factoextra)
library(lubridate)
```

```{r}
housing_sales<- read.csv("~/ADS/ADS 500B/Datasets/house_sales.csv") #reading in data
```

## Fixing the date column with lubridate

```{r}
housing_sales[2] <- housing_sales[2] %>%
  mutate(date = as.character(date)) %>%
  mutate(date = str_remove(date, "T000000")) %>%
  mutate(date = ymd(date))
housing_sales <- housing_sales %>%
  mutate(zipcode=as.factor(zipcode))

str(housing_sales)
```

## Looking at the state of the NAs

```{r}
house_na<- housing_sales %>%
  filter(is.na(housing_sales[,1]) | is.na(housing_sales[,2]) | is.na(housing_sales[,3]) | is.na(housing_sales[,4]) | is.na(housing_sales[,5]) | is.na(housing_sales[,6]) | is.na(housing_sales[,7]) | is.na(housing_sales[,8]) | is.na(housing_sales[,9]) | is.na(housing_sales[,10])| is.na(housing_sales[,11]) | is.na(housing_sales[,12]) | is.na(housing_sales[,13]) | is.na(housing_sales[,14]) | is.na(housing_sales[,15]) | is.na(housing_sales[,16]) | is.na(housing_sales[,17]) | is.na(housing_sales[,18]) | is.na(housing_sales[,19]) | is.na(housing_sales[,20]) | is.na(housing_sales[,21]))
```

```{r}
colSums(is.na(house_na))
```

## Data transformations we could explore

Make price into price_100, where each increase of unit denotes an increase in $100,000 dollars.

Turn waterfront, view, condition, and grade into ordered factors.


# Clean Data (WIP)
```{r}
housing_clean<- housing_sales %>% #will be updated as we decide what to do with NAs
  filter(price != is.na(price))
```

# Cleaning the Data

Removing all rows with NAs in them would result in losing 3995 rows being removed. This is 18.5% of the rows. There are a few variables that I think are nonnegotiable, and should have all rows with NAs removed. This list includes price.

```{r}
#Price breakdown

favstats(~price, data = housing_clean)

ggplot(data = housing_clean) +
  geom_histogram(aes(x = price))
```

Choosing how to clean the rest of the NAs out kind of relies on what variables we want to focus on. From the data I've seen, I think we could do a supervised learning with an equation that looks like the following:

price = intercept + var1 + var2 + var3 + var4 + etc.

I think I definitely want to include yr_built, zipcode, sqft_living, sqft_lot, bedrooms, and bathrooms in our exploratory data analysis. I would also suggest that we look at the condition, just to see how things line up.

I found the original dataset, which explains the sqft_living15 and sqft_lot15. The variable sqft_living15 is the square footage of interior housing living space for the nearest 15 neighbors. The variable sqft_lot15 is the square footage of the land lots of the nearest 15 neighbors. These were probably included as a sort of measure of wealth in the area.


```{r}
# Are sqft_living15 and sqft_living linearly related? Can one predict the other?
ggplot(data = housing_clean) +
  geom_point(aes(x = sqft_living15, y = sqft_living))

# Are sqft_lot15 and sqft_lot linearly related? Can one predict the other?
ggplot(data = housing_clean) +
  geom_point(aes(x = sqft_lot15, y = sqft_lot))
```


## Notes for Bedrooms (1134 NAs)

```{r}
ggplot(data = housing_clean) +
  geom_boxplot(aes(y=bedrooms))
```

- the row with 33 bedrooms should be smoothed (only sold for 640K)
- replace with median due to skew? or just remove rows with NAs

```{r}
housing_clean[15871,4]<- 3 #smooths the house that had 33 bedrooms
```

## Notes for Bathrooms (1068 NAs)

```{r}
ggplot(data = housing_clean) +
  geom_boxplot(aes(y=bathrooms))
```

- replace with median due to skew? or just remove rows with NAs


## Notes for Square Feet of Living Space (1110 NAs)

```{r}
ggplot(data = housing_clean) +
  geom_boxplot(aes(y=sqft_living))
```

- replace NA's with sqft_living15

```{r}
housing_clean <- housing_clean %>%
  mutate(sqft_living = ifelse(is.na(sqft_living), sqft_living15, sqft_living))
```


## Notes for Square Feet of Lot Space (1044)

```{r}
ggplot(data = housing_clean) +
  geom_boxplot(aes(y=sqft_lot))
```
- sqft lot > 1500000 -> noisy? only sold for $700,000 -> could replace with sqft_lot15

```{r}
housing_clean <- housing_clean %>%
  mutate(sqft_lot = ifelse(is.na(sqft_lot), sqft_lot15, sqft_lot))
```


# All the Cleaning In One Place
```{r}
# housing_clean<- housing_sales %>% #will be updated as we decide what to do with NAs
#   filter(price != is.na(price)) %>%
#   mutate(sqft_living = ifelse(is.na(sqft_living), sqft_living15, sqft_living)) %>%
#   mutate(sqft_lot = ifelse(is.na(sqft_lot), sqft_lot15, sqft_lot))

# housing_clean[15871,4]<- 3 #smooths the house that had 33 bedrooms

# colSums(is.na(housing_clean))
```


# Bedrooms

```{r}
housing_clean<- housing_clean %>%
  filter(id != 6306400140 & id != 1453602309 & id != 6896300380 & id != 2954400190 & id != 2569500210 & id != 2310060040 & id != 3374500520 & id != 7849202190 & id != 7849202299 & id != 9543000205 & id != 1222029077) #removes houses with zero bedrooms

housing_clean %>%
  arrange(bedrooms)

housing_clean<- housing_clean %>%
  mutate(bed_fact = as.factor(bedrooms),
         bath_char = as.character(bathrooms),
         bath_fact = fct_collapse(bath_char,
    "0 to 1" = c("0", "0.5", "0.75"), 
    "1.25 to 2" = c("1", "1.25", "1.5", "1.75", "2"), 
    "2.25-3" = c("2.25", "2.5", "2.75", "3") ,
    "3.25-4" = c("3.25", "3.5", "3.75", "4"),
    "4.25-5" = c("4.25", "4.5", "4.75", "5"),
    "5.25 and up" = c("5.25", "5.5", "5.75", "6", "6.25", "6.5", "6.75", "7.5", "7.75", "8")
  )) 

housing_clean %>%
  group_by(bed_fact) %>%
  summarise(min = min(sqft_living),
            max = max(sqft_living),
            median_sqft = median(sqft_living))

housing_clean %>%
  group_by(bed_fact) %>%
  summarise(n = n())

housing_clean<- housing_clean %>%
  mutate(bedrooms = case_when(
    is.na(bedrooms) & sqft_living < 1000 ~ 1,
    is.na(bedrooms) & sqft_living < 1500 ~ 2,
    is.na(bedrooms) & sqft_living < 2000 ~ 3,
    is.na(bedrooms) & sqft_living < 2600 ~ 4,
    is.na(bedrooms) & sqft_living < 2900 ~ 5,
    is.na(bedrooms) & sqft_living <= 3000 ~ 6,
    is.na(bedrooms) & sqft_living > 3000 ~ 7,
    .default = bedrooms
  ))
colSums(is.na(bed))

```

```{r}
housing_clean %>%
  group_by(bath_fact) %>%
  summarise(min = min(sqft_living),
            max = max(sqft_living),
            median_sqft = median(sqft_living))

housing_clean %>%
  group_by(bath_fact) %>%
  summarise(n = n())

housing_clean<- housing_clean %>%
  mutate(bathrooms = case_when(
    is.na(bathrooms) & sqft_living < 1000 ~ 1,
    is.na(bathrooms) & sqft_living < 1500 ~ 2,
    is.na(bathrooms) & sqft_living < 3000 ~ 3,
    is.na(bathrooms) & sqft_living < 4000 ~ 4,
    is.na(bathrooms) & sqft_living > 4000 ~ 5,
    .default = bathrooms
  ))
colSums(is.na(housing_clean))

housing_clean<- housing_clean %>%
  select(1:21)
```


```{r}
# housing_groups <- housing_clean %>%
#  mutate(bed_group = fct_collapse(bed_fact,
#    "one to three" = c("1", "2", "3"), # median sqft in the 1,000 - 2,000 range
#    "four to six" = c("4", "5", "6"), # median sqft in the 2,000 to 3,000 range
#    "seven to ten" = c("7", "8", "9", "10") # median sqft in the 3,000+ range
#  ))

# housing_groups %>%
#  group_by(bed_group) %>%
#  summarise(median = median(sqft_living))
```







## ZIPCODE (thinking of removing)

Is zipcode an important predictor of price? 

```{r, Price by zipcode}
housing_clean %>%
  group_by(zipcode) %>%
  summarise(mean = mean(price)) %>%
  ggplot() +
  geom_histogram((aes(x = mean))) +
  labs(title = "Mean Price by Zipcode")

housing_clean %>%
  group_by(zipcode) %>%
  summarise(mean = mean(price)) %>%
  arrange(desc(mean))
```

```{r, Top 3 mean price zip codes}
#Looking at the zipcodes with the top mean prices

housing_clean %>%
  filter(zipcode == "98039") %>%
  arrange(desc(price))

housing_clean %>%
  filter(zipcode == "98004") %>%
  arrange(desc(price))

housing_clean %>%
  filter(zipcode == "98040") %>%
  arrange(desc(price))
```

What about square footage?

```{r}
housing_clean %>%
  group_by(zipcode) %>%
  summarise(mean = mean(sqft_living15)) %>%
  ggplot() +
  geom_histogram((aes(x = mean))) +
  labs(title = "Mean Interior Square Footage by Zipcode")

housing_clean %>%
  group_by(zipcode) %>%
  summarise(mean = mean(sqft_living15)) %>%
  arrange(desc(mean))
```

