---
title: "ADS 500B Project: Cleaning and Initial Modeling"
author: "Madeline Chang"
date: "2024-03-01"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
library(dplyr)
library(tidyverse) 
library(tidymodels)
library(mosaic) 
library(cluster)  
library(factoextra)
library(lubridate)
```

```{r}
housing_sales<- read.csv("~/ADS/ADS 500B/Datasets/house_sales.csv") #reading in data
```

```{r}
housing_clean<- housing_sales %>% #will be updated as we decide what to do with NAs
  filter(price != is.na(price))

housing_clean[2] <- housing_clean[2] %>% 
  mutate(date = as.character(date)) %>%
  mutate(date = str_remove(date, "T000000")) %>%
  mutate(date = ymd(date)) #fixing the date with lubridate
```

```{r}
housing_clean[15871,4]<- 3

housing_clean <- housing_clean %>%
  mutate(price_10000 = price / 10000,
         sqft_living = ifelse(is.na(sqft_living), sqft_living15, sqft_living), #replaces NA living sqft with the mean of the nearest 15 neighbors' living sqft measurement
         sqft_living100 = sqft_living / 100,
         sqft_lot = ifelse(is.na(sqft_lot), sqft_lot15, sqft_lot), #replaces NA lot sqft with the mean of the nearest 15 neighbors' lot sqft measurement
         sqft_lot100 = sqft_lot / 100,
         sqft_above100 = sqft_above / 100,
         sqft_basement100 = sqft_basement / 100,
         bedrooms = case_when(
           is.na(bedrooms) & sqft_living < 1000 ~ 1,
           is.na(bedrooms) & sqft_living < 1500 ~ 2,
           is.na(bedrooms) & sqft_living < 2000 ~ 3,
           is.na(bedrooms) & sqft_living < 2600 ~ 4,
           is.na(bedrooms) & sqft_living < 2900 ~ 5,
           is.na(bedrooms) & sqft_living <= 3000 ~ 6,
           is.na(bedrooms) & sqft_living > 3000 ~ 7,
           .default = bedrooms
         ), #replaces NAs with approximations from the 500B_project_clean.RMD
         bathrooms = case_when(
           is.na(bathrooms) & sqft_living < 1000 ~ 1,
           is.na(bathrooms) & sqft_living < 1500 ~ 2,
           is.na(bathrooms) & sqft_living < 3000 ~ 3,
           is.na(bathrooms) & sqft_living < 4000 ~ 4,
           is.na(bathrooms) & sqft_living >= 4000 ~ 5,
           .default = bathrooms
         ), #replaces NAs with approximations from the 500B_project_clean.RMD
         zipcode=as.factor(zipcode),  #makes zipcode a factor
         condition_f = as.factor(condition),
         waterfront_f = as.factor(waterfront),
         view_f = as.factor(view),
         grade_f = as.factor(grade),
         renovated = ifelse(yr_renovated != 0, 1, 0)) %>%
  filter(bedrooms != 0) #removes houses with zero bedrooms

colSums(is.na(housing_clean))
```

```{r}
ggplot(data = housing_clean, aes(x=yr_built, y=price)) +
  geom_point() +
  geom_smooth(method = "lm") # perhaps don't use this?
```

```{r}
ggplot(data = housing_clean, aes(x=sqft_living, y=price)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
ggplot(data = housing_clean, aes(x=bedrooms, y=price)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
ggplot(data = housing_clean, aes(x=bathrooms, y=price)) +
  geom_point() +
  geom_smooth(method = "lm")
```

#Splitting test and train

```{r}
set.seed(123456)
housing_split <- initial_split(housing_clean, prop=0.5)
housing_train <- training(housing_split)
housing_test <- testing(housing_split)
```


# Building Linear Regression models with training data- PREDICTING PRICE

```{r}
model0<- lm(price ~ 1, data = housing_train)
summary(model0)
```

```{r}
model0_all<- lm(price ~ ., data = housing_train)
summary(model6)
```

```{r}
model1<- lm(price ~ sqft_living, data = housing_train)
summary(model1)
```

```{r}
model2<- lm(price ~ sqft_living + bedrooms + bathrooms, data = housing_train)
summary(model2)
```

```{r}
model3<- lm(price ~ sqft_living + grade, data = housing_train)
summary(model3)
```

```{r}
model4<- lm(price ~ condition + waterfront + view + grade, data = housing_train)
summary(model4)
```

```{r}
model5<- lm(price ~ condition + waterfront + view + grade + sqft_living + bedrooms, data = housing_train)
summary(model5)
```

```{r}
model6<- lm(price ~ date + floors + condition + waterfront + view + grade + sqft_living + sqft_lot + bedrooms + bathrooms + yr_built + sqft_above + sqft_basement, data = housing_train)
summary(model6)

```

With a Multiple $R^2$ of 0.6483, that means that 64.83% of the variability in this dataset is explained by this model (model6).

Model 6: Top 5 Variables with highest absolute value of t-value
  1. Grade: 38.776
  2. yr_built: -36.758
  3. waterfront: 26.013
  4. view: 13.388
  5. bedrooms:-12.779


Model 6 with transformed price, sqft_living, sqft_lot, sqft_above, and sqft_basement. This does NOT change any of the p-values, but it does make interpretation different.

Each increase in one unit of price = an increase of $10,000
Each increase in one unit of sqft_blank100 = an increase of 100 sqft

```{r}
model6_trans<- lm(price_10000 ~ date + floors + condition + waterfront + view + grade + sqft_living100 + sqft_lot100 + bedrooms + bathrooms + yr_built + sqft_above100 + sqft_basement100, data = housing_train)
summary(model6_trans)

```

What if we try to account for interactions?

```{r}
model7<- lm(price_10000 ~ date + floors + condition + waterfront + view + grade + sqft_living100 + sqft_lot100 + bedrooms + bathrooms + yr_built + sqft_above100 + sqft_basement100 + bedrooms:bathrooms + sqft_living100:sqft_lot100, data = housing_train)
summary(model7)
```

Adding interactions between bedrooms and bathrooms, along with sqft_living100 and sqft_lot100 has rendered sqft_lot100 not significant. It has increased the t-value of grade by 2.

The $R^2$ value is 0.6557, which means that 65.57% of the variability in this dataset can be explained by this model.

```{r}
model7a<- lm(price_10000 ~ date + floors + condition + waterfront + view + grade + sqft_living100 + bedrooms + bathrooms + yr_built + sqft_above100 + sqft_basement100 + bedrooms:bathrooms + sqft_living100:sqft_lot100, data = housing_train)
summary(model7a)
```

Removing the insignificant term does not change the R-Squared.

What if we removed the variables that are included in the interactions, and kept only the interactions?

```{r}
model8<- lm(price_10000 ~ date + floors + condition + waterfront + view + grade + yr_built + sqft_above100 + sqft_basement100 + bedrooms:bathrooms + sqft_living100:sqft_lot100, data = housing_train)
summary(model8)
```

Doing this renders the bedrooms and bathrooms interaction insignificant. I would suggest that we keep the separate bedrooms and bathrooms terms.

What if we kept the separate bedrooms and bathrooms terms, plus their interaction, but removed the interaction for sqft_living100 and sqft_lot100?

```{r}
model9<- lm(price_10000 ~ date + bedrooms + bathrooms + floors + condition + waterfront + view + grade + yr_built + sqft_above100 + sqft_basement100 + bedrooms:bathrooms , data = housing_train)
summary(model9)
```

Doing so gives us an $R^2$ value of 0.6544, which means that 65.44% of the variability in the dataset can be explained by this model (model9)



# A summary of the models created so far

Model                     $R^2$               
-------------------       --------------------   
model0                    NA                                             
model1                    0.481                 
model2                    0.4936                 
model3                    0.518                 
model4                    0.5258                 
model5                    0.5968                 
model6                    0.6483                 
model7                    0.6557
model7a                   0.6557
model8                    0.6421
model9                    0.6544

From these models, I will move forward with model7a, as it has the highest $R^2$ value, and less predictor variables than model7.

# Applying the model to the data
I'm going to basically take the model7a and apply it to the testing data using the function "augment". This will add some columns to the housing_test tibble, which provides the estimates based on model7a.

```{r}
fit_7a<- model7a %>%
  augment(housing_test)

fit_7a %>%
  rmse(price, .fitted) %>%
     pull(.estimate)
```


# Non-Linear regression models

## Choices for non-lm() regressions
  1. Regression tree


